---
title: "ML Assignment 4"
author: "Sean Bradford"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
rm(list=ls())

library(tidyverse)
library(ISLR)
library(factoextra)
library(flexclust)

pharm=read.csv('C:\\Users\\Sean\\OneDrive\\Desktop\\Grad School\\Machine Learning\\Module 6 - K-mean Clustering Algorithm\\Pharmaceuticals.csv')
head(pharm)
```

A. Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in 
conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) 
used, the number of clusters formed, and so on.

```{r}
set.seed(111)
# Retain only quantitative variables from original df
dfp=data.frame(pharm[,c(3:11)])
summary(dfp)
```

```{r}
# Scaling the data (z-score)
dfp=scale(dfp)
dis=get_dist(dfp)
fviz_dist(dis)
```

```{r}
# Determining k 
fviz_nbclust(dfp,kmeans,"wss")
```

```{r}
fviz_nbclust(dfp,kmeans,"silhouette")
```

Both charts indicate that 5 is the ideal number of clusters. 

```{r}
# k means clustering (Euclidean distance)
k.5=kmeans(dfp,centers = 5, nstart=25)
k.5
k.5$centers

```

```{r}
k.5$size
```

```{r}
k.5$cluster[6]
```

```{r}


fviz_cluster(k.5,data=dfp)
```


```{r}
# k means clustering using Manhattan distance for comparison
set.seed(111)
kc.5=kcca(dfp, k=5, kccaFamily("kmedians"))
kc.5
```

```{r}
clusters_index=predict(kc.5)
dist(kc.5@centers)
image(kc.5)
points(dfp,col=clusters_index,pch=19)
```

B. Interpret the clusters with respect to the numerical variables used in forming the clusters.

```{r}
# Centroids for Euclidean distance model
k.5$centers
```

```{r}
# Create profile plot of cluster centroids

dfp1=data.frame(pharm[,c(3:11)])

plot(c(0),xaxt='n',ylab="", type="l",
     ylim=c(min(k.5$centers),max(k.5$centers)),xlim=c(0,9))
axis(1,at=seq(1:9),labels = names(dfp1),cex.axis=0.55,las=3)

for (i in c(1:5))
  lines(k.5$centers[i,],lty=i,lwd=2,col= if(i %in% 1) {
    col="blue"
  } else if(i %in% 2) {
    col="green"
  } else if(i %in% 3) {
    col="purple"
  } else if(i %in% 4) {
    col="orange"
  } else {
    col="red"
  }
  )

text(x=1,y=k.5$centers[,1],labels=paste("Cluster",c(1:5)))
```
Characteristics of each cluster:

Cluster 1 (blue):   Lower Beta; Higher Net Proft Margin
Cluster 2 (green):  Lower Market Cap and ROA; Higher Beta and Leverage
Cluster 3 (purple): Higher P/E Ratio; Lower ROE and Net Profit Margin
Cluster 4 (orange): Higher Market Cap, ROE, ROA & Asset Turnover; Lower leverage
Cluster 5 (red):    Higher Revenue Growth; Lower Asset Turnover


C. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in 
forming the clusters)

```{r}
# Seperate clusters to analyze unused variables
c1=pharm[c(1,4,7,10,21,16,19,3),c(12:14)]
c2=pharm[c(12,8,6),c(12:14)]
c3=pharm[c(2,18),c(12:14)]
c4=pharm[c(11,17,13,15),c(12:14)]
c5=pharm[c(5,20,14,9),c(12:14)]
```

```{r}
c1
c2
c3
c4
c5
```
For Median Recommendation variable:

Cluster 1: Mainly Hold 
Cluster 2: Mainly Hold 
Cluster 3: 1 mod buy and 1 hold
Cluster 4: 2 mod buy and 2 hold
Cluster 5: 2 mod buy and 2 mod sell

Clusters 1 and 2 have similar Revenue Growth and P/E Ratio. There appears to be a pattern between those two factors and receiving a Hold recommendation.

Clusters 3-5 are top 3 in Revenue Growth and half the recommendations for each cluster are Moderate Buy. Pattern showing that pharm companies with higher revenue growth are more likely to receive buy recommendations.


D. Provide an appropriate name for each cluster using any or all of the variables in the dataset.

Cluster 1: High Net Profit Margin
Cluster 2: High Beta & Leverage; Low ROA & Market Cap
Cluster 3: Very High P/E Ratio; Very low Net Profit Margin
Cluster 4: High ROE,ROA,Asset Turnover; Low Leverage
Cluster 5: High Revenue Growth; Very Low Asset Turnover